---
title: "PARE-Seq Analyses 2022"
author: "Scarlet Bliss"
date: "1/26/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(readr)
library(car)
library(MASS)
library(multilevel)
library(sjPlot)
library(glmnet)
library(olsrr)
library(effsize)
library(gridExtra)
```

```{r}
df6 <- read.csv("pareseq_full_12.29.22.csv")
```

#Recoding to make binary variables for analysis 
```{r}
df7 <- df6 %>% mutate(racecat2=dplyr::recode(racecat, #making a binary race/ethnicity variable
                                             "0"=0, #white
                                             "1"=1, #black
                                             "3"=0, #asian
                                             "2"=1, #middle eastern/north african
                                             "4"=1, #indigenous
                                             "6"=1))  %>% #biracial/multiracial
              mutate(workhrs2=dplyr::recode(workhrs, #making a binary work hours variable
                                            "0"=0,#no
                                            "1"=0, #1-10
                                            "2"=1, #11-20
                                            "3"=1, #21-30
                                            "4"=1)) %>% #31+
              mutate(gender=dplyr::recode(gender,
                                          "0"=1, #females
                                          "1"=0)) %>% #males
             mutate(workspace2 = case_when(workspace == 2 ~0, #making a binary workspace rating variable
                                           workspace==0~0, #high quality
                                            workspace==3 ~1)) #low quality
df7$school <- factor(df7$school, levels=c("RU", "QU", "DTCC", "NSCC"))
df7 <- df7 %>% mutate_at(c(5:9,14,17), as.factor)

```


#Summary statistics: Mean, SD, T-Tests, Cohen's d
```{r}
#learning gains: mean and SD
mean(df7$totalpre)
mean(df7$totalpost)
sd(df7$totalpre)
sd(df7$totalpost)

TTest.lg <- t.test(df7$totalpost, df7$totalpre, 
                paired = TRUE, alternative = "greater")
TTest.lg
#effect size
cohen.d(df7$totalpost, df7$totalpre) #0.221

#stem identity gains: mean and SD
mean(df7$stemidenpre)
mean(df7$stemidenpost)
sd(df7$stemidenpre)
sd(df7$stemidenpost)

TTest.si <- t.test(df7$stemidenpost, df7$stemidenpre, 
                paired = TRUE, alternative = "greater")
TTest.si
#effect size
cohen.d(df7$stemidenpost, df7$stemidenpre) #0.299, small
```

ICCs: influence of school as cluster variable
```{r}
aov.lg <- aov(totaldiff ~ school, df7)
aov.st <- aov(stemIDchange ~ school, df7)

ICC1(aov.lg)
ICC1(aov.st)

#The ICC here for learning gains is 0.08, but in the full mixed effects model it was 0.12. Which should be reported?
```


#LEARNING GAINS: full model
```{r}
lg_full <- lm(totaldiff ~ gender+racecat2+parentedu+workhrs+workspace2+completion+school, data=df7)
summary(lg_full)
Anova(lg_full)

sjPlot::tab_model(lg_full,
                  show.re.var=TRUE,
                  show.se=TRUE,
                  dv.labels= 'Full Model: Learning Gains by All Predictors')
```
Assessing collinearity of covariates
```{r}
vcov(lg_full)
ols_vif_tol(lg_full)
#none seem to be highly collinear
```


#LEARNING GAINS Bivariate analyses
Learning gains & Racecat2 (binary) bivariate model
```{r}
lg_raceeth_bin <- lm(totaldiff ~ racecat2, data=df7)
summary(lg_raceeth_bin)
Anova(lg_raceeth_bin)

sjPlot::tab_model(lg_raceeth_bin,
                  show.re.var=TRUE,
                  dv.labels= 'Bivariate Analyses: Race/Ethnicity')

#p = 0.087
```
Learning gains & racecat (categorical race variable)
```{r}
lg_raceeth <- lm(totaldiff ~ racecat, data=df7)
summary(lg_raceeth)
Anova(lg_raceeth)

sjPlot::tab_model(lg_raceeth,
                  show.re.var=TRUE,
                  dv.labels= 'Bivariate Analyses: Race/Ethnicity')
#indigenous and bi-racial/multiracial categories statistically significant. 
```

Learning gains & Workhrs Bivariate model
```{r}
lg_workhrs <- lm(totaldiff ~ workhrs, data=df7)
summary(lg_workhrs)
Anova(lg_workhrs)

sjPlot::tab_model(lg_workhrs,
                  show.re.var=TRUE,
                  dv.labels= 'Bivariate Analyses: workhrs')

#no statistical significance (category 4, 31+h, approaches significance. p=0.08)
```

Learning gains & Gender Bivariate model
```{r}
lg_gender <- lm(totaldiff ~ gender, data=df7)
summary(lg_gender)
Anova(lg_gender)

sjPlot::tab_model(lg_gender,
                  show.re.var=TRUE,
                  dv.labels= 'Bivariate Analyses: Gender')
```

Learning gains & Workspace bivariate model
```{r}
lg_workspace <- lm(totaldiff ~ workspace2, data=df7)
summary(lg_workspace)
Anova(lg_workspace)

sjPlot::tab_model(lg_workspace,
                  show.re.var=TRUE,
                  dv.labels= 'Bivariate Analyses: Self-rated quality of workspace')
#approaches significance, p=0.076
```


Learning gains & Parentedu bivariate model
```{r}
lg_parentedu <- lm(totaldiff ~ parentedu, data=df7)
summary(lg_parentedu)
Anova(lg_parentedu)

sjPlot::tab_model(lg_parentedu,
                  show.re.var=TRUE,
                  dv.labels= 'Bivariate Analyses: Parent Education')
```

Learning gains & Completion bivariate model
```{r}
lg_completion <- lm(totaldiff ~ completion, data=df7)
summary(lg_completion)
Anova(lg_completion)

sjPlot::tab_model(lg_completion,
                  show.re.var=TRUE,
                  dv.labels= 'Bivariate Analyses: Completion')
```
Learning gains & school bivariate model
```{r}
lg_school <- lm(totaldiff ~ school, data=df7)
summary(lg_school)
Anova(lg_school)

sjPlot::tab_model(lg_school,
                  show.re.var=TRUE,
                  dv.labels= 'Bivariate Analyses: School')
```
Table of bivariate results: learning gains
```{r}
learning.gains.results <- rbind(
  tidy(lg_gender, conf.int=T)[2,],
  tidy(lg_parentedu, conf.int=T)[2,],
  tidy(lg_raceeth_bin, conf.int=T)[2,],
  tidy(lg_workhrs, conf.int=T)[2,],
  tidy(lg_workhrs, conf.int=T)[3,],
  tidy(lg_workhrs, conf.int=T)[4,],
  tidy(st_workhrs, conf.int=T)[5,],
  tidy(lg_workspace, conf.int=T)[2,],
  tidy(lg_completion, conf.int=T)[2,],
  tidy(lg_completion, conf.int=T)[3,],
  tidy(lg_completion, conf.int=T)[4,],
  tidy(lg_school, conf.int=T)[2,],
  tidy(lg_school, conf.int=T)[3,],
  tidy(lg_school, conf.int=T)[4,]) 

learning.gains.results
```


#STEM IDENTITY:full fixed-effects model
```{r}
st_full <- lm(stemIDchange ~ gender+racecat2+workhrs+parentedu+completion+workspace2+school, data=df7)
summary(st_full)
Anova(st_full)

sjPlot::tab_model(st_full,
                  show.re.var=TRUE,
                  show.se=TRUE,
                  dv.labels= 'Full Fixed Effects Model: STEM Identity by All Predictors')
```

#BIVARIATE STEM ID MODELS
STEM ID & Race/ethnicity (binary)
```{r}
st_raceeth_bin <- lm(stemIDchange ~ racecat2, data=df7)
summary(st_raceeth_bin)
Anova(st_raceeth_bin)

sjPlot::tab_model(st_raceeth_bin,
                  show.re.var=TRUE,
                  dv.labels= 'STEM ID Bivariate Analyses: Race/Ethnicity')
```

STEM ID & Gender
```{r}
st_gender <- lm(stemIDchange ~ gender, data=df7)
summary(st_gender)
Anova(st_gender)

sjPlot::tab_model(st_gender,
                  show.re.var=TRUE,
                  dv.labels= 'STEM ID Bivariate Analyses: Gender')
```

STEM ID & Parentedu
```{r}
st_parentedu <- lm(stemIDchange ~ parentedu, data=df7)
summary(st_parentedu)
Anova(st_parentedu)

sjPlot::tab_model(st_parentedu,
                  show.re.var=TRUE,
                  dv.labels= 'STEM ID Bivariate Analyses: Parent Education')
```

STEM ID & Workhrs
```{r}
st_workhrs <- lm(stemIDchange ~ workhrs, data=df7)
summary(st_workhrs)
Anova(st_workhrs)

sjPlot::tab_model(st_workhrs,
                  show.re.var=TRUE,
                  dv.labels= 'STEM ID Bivariate Analyses: Work Hours')
```

STEM ID & Workspace
```{r}
st_workspace <- lm(stemIDchange ~ workspace, data=df7)
summary(st_workspace)
Anova(st_workspace)

sjPlot::tab_model(st_workspace,
                  show.re.var=TRUE,
                  dv.labels= 'STEM ID Bivariate Analyses: Work Space')
```
STEM ID & Completion
```{r}
st_completion <- lm(stemIDchange ~ completion, data=df7)
summary(st_completion)
Anova(st_completion)

sjPlot::tab_model(st_completion,
                  show.re.var=TRUE,
                  dv.labels= 'STEM ID Bivariate Analyses: Completion')
```
STEM ID & School
```{r}
st_school <- lm(stemIDchange ~ school, data=df7)
summary(st_school)
Anova(st_school)

sjPlot::tab_model(st_school,
                  show.re.var=TRUE,
                  dv.labels= 'STEM ID Bivariate Analyses: Completion')
```
Table of bivariate results: STEM Idenity
```{r}
STEM.identity.results <- rbind(
  tidy(st_gender, conf.int=T)[2,],
  tidy(st_parentedu, conf.int=T)[2,],
  tidy(st_raceeth_bin, conf.int=T)[2,],
  tidy(st_workhrs, conf.int=T)[2,],
  tidy(st_workhrs, conf.int=T)[3,],
  tidy(st_workhrs, conf.int=T)[4,],
  tidy(st_workhrs, conf.int=T)[5,],
  tidy(st_workspace, conf.int=T)[2,],
  tidy(st_completion, conf.int=T)[2,],
  tidy(st_completion, conf.int=T)[3,],
  tidy(st_completion, conf.int=T)[4,],
  tidy(st_school, conf.int=T)[2,],
  tidy(st_school, conf.int=T)[3,],
  tidy(st_school, conf.int=T)[4,]) 

STEM.identity.results
```


#DESCRIPTIVE STATISTICS (Tables 4a & 4b)

Learning gains (4a)
Learning Gains by gender
```{r}
tapply(df7$totalpre, df7$gender, mean)
tapply(df7$totalpost, df7$gender, mean)
```
Learning Gains by Work hours
```{r}
tapply(df7$totalpre, df7$workhrs, mean)
tapply(df7$totalpost, df7$workhrs, mean)
```

Learning Gains by Parent edu
```{r}
tapply(df7$totalpre, df7$parentedu, mean)
tapply(df7$totalpost, df7$parentedu, mean)
```


Learning Gains by Workspace
```{r}
tapply(df7$totalpre, df7$workspace2, mean)
tapply(df7$totalpost, df7$workspace2, mean)

```


Learning Gains by school
```{r}
tapply(df7$totalpre, df7$school, mean)
tapply(df7$totalpost, df7$school, mean)
```

STEM Identity Change (4b)
STEM Identity Change by Race/Ethnicity
```{r}
tapply(df7$stemidenpre, df7$racecat2, mean)
tapply(df7$stemidenpost, df7$racecat2, mean)
```

STEM ID by gender
```{r}
tapply(df7$stemidenpre, df7$gender, mean)
tapply(df7$stemidenpost, df7$gender, mean)
```
STEM ID by Work hours
```{r}
tapply(df7$stemidenpre, df7$workhrs, mean)
tapply(df7$stemidenpost, df7$workhrs, mean)
```

STEM ID by Parent edu
```{r}
tapply(df7$stemidenpre, df7$parentedu, mean)
tapply(df7$stemidenpost, df7$parentedu, mean)
```


STEM ID by Workspace
```{r}
tapply(df7$stemidenpre, df7$workspace2, mean)
tapply(df7$stemidenpost, df7$workspace2, mean)
```

STEM ID by School
```{r}
tapply(df7$stemidenpre, df7$school, mean)
tapply(df7$stemidenpost, df7$school, mean)
```

Learning Gains by Race/Ethnicity
```{r}
tapply(df7$totalpre, df7$racecat2, mean)
tapply(df7$totalpost, df7$racecat2, mean)
```


Work hours and race/ethnicity crosstab
```{r}
CrossTable(df7$racecat2, df7$workhrs)
CrossTable(df7$racecat2, df7$workhrs2)
```

#Box Plots (Fig. 4)
```{r}
df8 <- df7%>% pivot_longer(cols=c('totalpre', 'totalpost', 'stemidenpre', 'stemidenpost'),
                           names_to='score_name',
                           values_to='value')
#learning gains
d.boxplot.lg <- df8%>%filter(score_name==c('totalpre', 'totalpost'))
d.boxplot.lg$score_name <- factor(d.boxplot.lg$score_name, levels=c('totalpre', 'totalpost'))
boxplot.lg <- ggplot(data=d.boxplot.lg, aes(x=score_name, y=value))+
  geom_boxplot(width=0.35)+
  theme_classic()+
  labs(y="Assessment Score (out of 10)")+ 
  scale_x_discrete(labels=c('Pre-Assessment', 'Post-Assessment'))+
  theme(axis.title.x=element_blank())
boxplot.lg

#stem identity
d.boxplot.st <- df8%>%filter(score_name==c('stemidenpre', 'stemidenpost'))
d.boxplot.st$score_name <- factor(d.boxplot.st$score_name, levels=c('stemidenpre', 'stemidenpost'))
boxplot.st <- ggplot(data=d.boxplot.st, aes(x=score_name, y=value))+
  geom_boxplot(width=0.35)+
  theme_classic()+
  labs(y="STEM Identity Score (out of 7)")+ 
  scale_x_discrete(labels=c('Pre-Survey', 'Post-Survey'))+
  theme(axis.title.x=element_blank())
boxplot.st

p <- grid.arrange(boxplot.lg, boxplot.st, ncol=2) 
ggsave("outcome.boxplots.png", p, height = 4, width = 7)
```

